Identifying covariate traits of trance and possession phenomena in the Ethnographic

### 2022-10-29 Readme.Rmd
now is good

This is the online supplement for the paper "Possession trance co-varies with measures of social rigidity in the Ethnographic Atlas 2022". The current version of the paper is [here](https://docs.google.com/document/d/1qbcsCeQ-Tw2W-yhQvvYd1Nk5ImQznqWBmiHX09lB_24/edit?usp=sharing).

@@

Our informal priors are that trance and possession phenomena result from complex interchanges between indidvidual psychopathology and social organisation, and that, even if we can recognise broader patterns, the effects of social rigidity and marginalisation will be idiosyncratic on the ground. This means that both the outcome (trance and possession) and the predictors (e.g. measures of social rigidity) will be much more loosely defined than in similar analyses. So we expect any effects we find to be relatively modest.

@@

The paper says that trance induced by possession is not a random, idiosyncratic pathology but a recurrent psycho-social trait that is at least partly motivated by other social factors. This is recurrent claim in the literature. Here, we revisit the main points of this argument.

@@

@@

```{r }
ld %>% 
  filter(var_title != 'Trance states') %>% 
  distinct(var_id,var_title,var_definition) %>%
  arrange(var_id) %>% 
  rename(ID = var_id, Name = var_title, Definition = var_definition) %>% 
  kable('simple')
```

@@

We use *hyperparameter tuning* to find the best possible model.  The model is fit on a training set of `length(unique(sccs$soc_id))` societies with information on trance and possession in the standard cross-cultural sample. The hyperparameter grid is printed in the Appendix.

@@

The overall results go some way in confirming hypotheses on the cross-cultural correlates of possession trance in the literature. @bourguignon1968 explicitly tests the variables of slavery, jurisdictional hierarchy, settlement patterns, and kin groups. @shaara1992 point to a high value placed on female virginity at marriage, captured here by norms of premarital sexual behaviour for girls, @wood2018complexity point to the role of moralising high gods. These are all important variables in our cross-cultural model. @greenbaum1973 and @shaara1992 point to sexual division of labour but this is not relevant in our model.

>>>>

# Online supplement: Possession trance co-varies with measures of social rigidity in the Ethnographic Atlas

This is the online supplement for the paper “Possession trance co-varies
with measures of social rigidity in the Ethnographic Atlas 2022”. The
current version of the paper is
[here](https://docs.google.com/document/d/1qbcsCeQ-Tw2W-yhQvvYd1Nk5ImQznqWBmiHX09lB_24/edit?usp=sharing). The `helper` file creates the data files. The `model` files run the Gradient Boosting Model and the Bayestraits model, respectively.

@@

Ee expect any effects we find to be relatively modest. This is because trance and possession phenomena result from complex interchanges between individual psychopathology and social organisation, and that, even if we can recognise broader patterns, the effects of social rigidity and marginalisation will be idiosyncratic on the ground. This means that both the outcome (trance and possession) and the predictors (e.g. measures of social rigidity) will be much more inherently variable than in similar analyses.

@@

The paper says that trance induced by possession is not a random, idiosyncratic pathology but a recurrent psycho-social trait that is at least partly motivated by other social factors. This is recurrent claim in the literature. Here, we revisit the main points of this claim.

@@

Broadly speaking, we have measures of:

1. Social complexity (including monotheism, which we assume to be more prevalent in more complex societies, see @botero2014ecology)
2. Social rigidity (including cognatic kinship, which @shaara1992 uses as an example of a more flexible social blueprint)
3. Social participation, with an emphasis on gender roles in social activites, including work.

@@

Variables 8, 30:33, 66, 68, 70, 72, 73 were lifted from previous work, most notably @bourguignon1968 and @greenbaum1973. In addition, Variables 15, 23, 27, 74, 113 were taken as approximations for social rigidity, while variables 53, 54, 78 for participation, based on the literature above.

```{r eval = F}
pred_table = ld %>% 
  filter(var_title != 'Trance states') %>% 
  distinct(var_id,var_title,var_definition) %>%
  arrange(var_id) %>% 
  rename(ID = var_id, Name = var_title, Definition = var_definition)

write_tsv(pred_table, 'data/pred_table_temp.tsv')
```

```{r }
pred_table = read_tsv('data/pred_table_added.tsv')
pred_table %>% 
  kable('simple')
```

@@

We use *hyperparameter tuning* to find the best possible model.  The model is fit on a training set of `r length(unique(sccs$soc_id))` societies with information on trance and possession in the standard cross-cultural sample. The hyperparameter grid is printed in the Appendix.

@@

The overall results go some way in confirming hypotheses on the cross-cultural correlates of possession trance in the literature. @bourguignon1968 explicitly tests the variables of domestic organisation slavery, jurisdictional hierarchy, settlement patterns, and kin groups. @shaara1992 point to a high value placed on female virginity at marriage, captured here by norms of premarital sexual behaviour for girls, @wood2018complexity point to the role of moralising high gods. These are all important variables in our cross-cultural model. @greenbaum1973 and @shaara1992 point to sexual division of labour but this is not relevant in our model.
